From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-06-06 paper.tex
Brayns gallery

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, OpenGL-like local illumination for
typical data sets used with rasterization (up to hundred million triangles) can
be done at similar framerates to OpenGL~\cite{ospray}. For small datasets,
OpenGL performs better, but for larger data sets raytracing outperforms OpenGL.
This is due to a better scalability with respect to the data set size
($O(\log{N})$ vs $O(n)$). Furthermore, CPU-based raytracing allows the rendering
of larger data sets without any level-of-detail algorithms due to the larger
memory size. Last, but not least, advanced rendering algorithms such as
shadowing, reflections and global illuminations are significantly easier to
implement in a raytracer. The only area where rasterization provides a benefit
is for rendering at very high frame rates, needed for example for immersive
visualization, or very high resolutions.

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. This application called Brayns can load and visualize our
data sets in a variety of modes and integrates with our messaging solution to be
accessible from Python. We are currently integrating the key algorithms from
Livre for out-of-core volume rendering in Brayns, which will also facilitate
mixing polygonal with large volumetric data in a single scene.

\FIXME{teaser gallery}

>>>>

Advances in computer hardware and software have brought raytracing to the point
where it replaces classical rasterization for virtually all use cases in
scientific visualization. On one hand, OpenGL-like local illumination for
typical data sets used with rasterization (up to hundred million triangles) can
be done at similar framerates to OpenGL~\cite{ospray}. For small datasets,
OpenGL performs better, but for larger data sets raytracing outperforms OpenGL.
This is due to a better scalability with respect to the data set size
($O(\log{N})$ vs $O(n)$). Furthermore, CPU-based raytracing allows the rendering
of larger data sets without any level-of-detail algorithms due to the larger
memory size. Last, but not least, advanced rendering algorithms such as
shadowing, reflections and global illumination (\fig{fBrayns}, left) are
significantly easier to implement in a raytracer. The only area where
rasterization provides a benefit is for rendering at very high frame rates,
needed for example for immersive visualization, or very high resolutions.

\begin{figure}[ht]\center
  \includegraphics[height=5cm]{images/Brayns003}\hfil
  \includegraphics[height=5cm]{images/Brayns001}
  \caption{\label{fBrayns}Interactive raytracing for neuroscience}
\end{figure}

It is for these reasons that interactive raytracing is the future technology for
interactive and offline rendering at Blue Brain. We have developed a first open
source implementation of a visualization engine with different backends:
OSPRay~\cite{ospray} for CPU-based raytracing and OptiX~\cite{optix} for
GPU-based raytracing. This application called Brayns can load and visualize our
data sets in a variety of modes and integrates with our messaging solution to be
accessible from Python. We are currently integrating the key algorithms from
Livre for out-of-core volume rendering in Brayns, which will also facilitate
mixing polygonal with large volumetric data in a single scene (\fig{fBrayns},
right).
